Carryover effects in free recall reveal how prior experiences influence memories of new experiences

### 2023-01-10 paper/main.tex
backing up methods additions...

Feature clustering describes a participant’s tendency to recall similar
presented items together in their recall sequences, where ``similarity''
considers one given feature dimension (e.g., category, color, etc.). Computing
the clustering score for one feature dimension starts by considering those
feature values from the first word the participant recalled on the list. Next,
we sort all not-yet-recalled words in ascending order according to their
feature-based distance to the just-recalled item (see \textit{Defining
feature-based distances}). We then compute the percentile rank of the observed
next recall. We averaged these percentile ranks across all of the participant’s
recalls for the current list to obtain a single uncorrected clustering score
for the list, for the given feature dimension. We repeat this process for each
feature dimension in turn to obtain a single uncorrected clustering score for
each list, for each feature dimension.

@@

>>>>

Feature clustering scores describe participants' tendencies to recall similar
presented items together in their recall sequences, where ``similarity''
considers one given feature dimension (e.g., category, color, etc.). We base
our main approach to computing clustering scores on analogous temporal and semantic clustering scores
developed by~\citep{PolyEtal09}. Computing the clustering score for one feature
dimension starts by considering those feature values from the first word the
participant recalled on the list. Next, we sort all not-yet-recalled words in
ascending order according to their feature-based distance to the just-recalled
item (see \textit{Defining feature-based distances}). We then compute the
percentile rank of the observed next recall. We averaged these percentile ranks
across all of the participant’s recalls for the current list to obtain a single
uncorrected clustering score for the list, for the given feature dimension. We
repeat this process for each feature dimension in turn to obtain a single
uncorrected clustering score for each list, for each feature dimension.

@@

Suppose that two lists contain unequal numbers of items of each size. For
example, suppose that list $A$ contains all ``large'' items, whereas list $B$
contains an equal mix of ``large'' and ``small'' items. For a participant
recalling list $A$, any correctly recalled item will necessarily match the size
of the previous correctly recalled item. In other words, successively recalling
several list $A$ items of the same size is essentially meaningless, since
\textit{any} correctly recalled list $A$ word will be large. In contrast,
successively recalling several list $B$ items \textit{could} be meaningful,
since (early in the recall sequence) the yet-to-be-recalled items come from a
mix of sizes. However, once all of the ``small'' items on list $B$ have been
recalled, the best possible next matching recall will be a large item. And all
subsequent correct recalls must also be large items-- so for those later
recalls it becomes difficult to determine whether the participant is
successively recalling ``large'' items because they are organizing their
memories according to size, or (alternatively), whether they are simply
recalling the yet-to-be-recalled items in a random order. In general, the
precise order and blend of feature values expressed in a given list, the
orders and numbers of correct recalls a participant makes, the number of itervening
presentation positions between successive recalls, and so on, can all affect the
range of clustering scores that are possible to observe for a given list.  The uncorrected
clustering score therefore conflates participants' actual memory organization with
other ``nuisance'' factors.

Following our prior work~\citep{HeusEtal17}, we used a permutation-correction
procedure to help isolate the behavioral aspects of clustering that we were
most interested in. After computing the uncorrected clustering score (for the
given list and observed recall sequence), we compute a ``null'' distribution of
$n$ additional clustering scores after randomly shuffling the recall order (we
use $n = 500$ in the present study). This null distribution represents an
approximation of the range of clustering scores one might expect to observe by
``chance,'' given that a hypothetical participant was \textit{not} truly
clustering their recalls, but where the hypothetical participant studied and
recalled exactly the same items (with the same features) as the true
participant.  We define the permutation-corrected clustering score as the percentile rank of the
observed uncorrected clustering score in this estimated null distribution.  In this way,
a corrected score of 1 indicates that the observed score was greater than any clustering score
one might expect by chance; in other words, good evidence that the participant was truly clustering their
recalls along the given feature dimension.
